KidsAIR home activity and characteristics cleaning

### 2020-04-24 exposure_models.Rmd
exposure analysis updates

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(emmeans)
library(broom)
library(MuMIn)
library(leaps)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(area == "AK") %>% 
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home, .keep_all = TRUE) %>% 
  select(home, area, sampling_day, starts_with("home_act"))

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  left_join(sums_temp_change, by = c("area", "home")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home")) %>% 
  select(pm_mean_sampling_period, area, income_3level, home_act_windows, 
                 stove_age_3level, home_windows_2level,
                 chimney_clean_3level, wood_collect_2level, 
                 home_act_door, home_act_smoking, home_act_sweep,
                 stove_grade_3level, moisture_closest_2level,
                 total_residents_2level, sums_events_sampling_period, mean_temp_sampling_period_5) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  na.exclude()
```

@@

Looking at Lasso and Ridge regression
May be beyond the scope of this paper, but keep code and references:

@@

>>>>

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(emmeans)
library(broom)
library(MuMIn)
library(leaps)
library(glmnet)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

@@

####### Format ambient data to merge with analysis data below
ambient_data <- exposure_analysis_data_long %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  #filter(area == "WMT") %>%
  filter(!is.na(sampling_day)) %>% 
  filter(!is.na(sample_date)) %>%
  select(area, home, child_id_char, sample_date, sampling_day, pm_mean_daily,
         zip:mean_temp_roll_4day) %>% 
  group_by(home) %>% 
  mutate(amb_pm_sampling_period_5 = mean(amb_pm_24hr, na.rm = TRUE)/5,
         mean_temp_sampling_period_5 = mean(mean_temp, na.rm = TRUE)/5) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup()

####### Format home activity data to merge with analysis data below
home_activity_data <- exposure_analysis_data_medium %>% 
  filter(treatment == "Placebo") %>% 
  #filter(winter_id == 1) %>% 
  select(home, area, home_winter_id, sampling_day, starts_with("home_act")) %>% 
  group_by(area, home, home_winter_id) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id) %>% 
  mutate_at(vars(starts_with("home_act")), as.numeric) %>% 
  mutate_at(vars(starts_with("home_act")), funs(. - 1)) %>% 
  mutate_at(vars(starts_with("home_act")), funs(sum(.))) %>% 
  ungroup() %>% 
  distinct(area, home,  home_winter_id, .keep_all = TRUE)

####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  filter(area == "WMT") %>% 
  group_by(area, home) %>% 
  distinct(home, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  group_by(home) %>% 
  mutate(residents_smoke = if_else(winter_id == 1 & is.na(residents_smoke),
                                   lead(residents_smoke), residents_smoke)) %>% 
  ungroup() %>% 
  left_join(sums_temp_change, by = c("area", "home", "home_winter_id")) %>% 
  left_join(ambient_data, by = c("area", "home")) %>% 
  select(-starts_with("home_act")) %>% 
  left_join(home_activity_data, by = c("area", "home", "home_winter_id")) %>% 
  select(pm_mean_sampling_period, area, income_3level, 
                 stove_age_3level, home_windows_2level,
                 chimney_clean_3level, wood_collect_2level, 
                 home_act_door, home_act_smoking, home_act_sweep, home_act_windows, 
                 stove_grade_3level, moisture_closest_2level,
                 total_residents_2level, mean_temp_sampling_period_5) %>% 
  filter(!is.na(pm_mean_sampling_period))

dep_var <- analysis_data$pm_mean_sampling_period

ind_vars <- analysis_data %>% 
  select(-pm_mean_sampling_period, -area) %>% 
  data.matrix()
```

@@

Looking at Lasso regression
May be beyond the scope of this paper, but keep code and references:

@@

Looking at Ridge regression:

https://drsimonj.svbtle.com/ridge-regression-with-glmnet

```{r}
ridge_results <- cv.glmnet(x = ind_vars, y = dep_var, alpha = 0)

plot(ridge_results)

opt_lambda <- ridge_results$lambda.min
opt_lambda

fit <- ridge_results$glmnet.fit
summary(fit)

y_predicted <- predict(fit, s = opt_lambda, newx = ind_vars)

# Sum of Squares Total and Error
sst <- sum((dep_var - mean(dep_var))^2)
sse <- sum((y_predicted - dep_var)^2)

# R squared
rsq <- 1 - sse / sst
rsq
```

# Train and test datasets

```{r}
# Random sample indexes
train_index <- sample(1:nrow(adult), 0.8 * nrow(adult))
test_index <- setdiff(1:nrow(adult), train_index)

dep_var <- analysis_data$pm_mean_sampling_period

ind_vars <- analysis_data %>% 
  select(-pm_mean_sampling_period, -area) %>% 
  data.matrix()
```

### 2020-04-24 home_char_act_cleaning.Rmd
exposure analysis updates

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)
```

@@

# Load, format, and save home activity data
```{r}
##### WMT #####
# Load data
home_act_initial_wmt <- read_csv("Input/WMT/wmt_home_activity_20200113.csv")

@@

# Save as RDS
# write_rds(wmt_home_act, "Output/wmt_home_act.rds")

##### NN #####
# Load data
home_act_initial_nn <- read_xlsx("Input/NN/nn_home_activity_20200113.xlsx")

@@

# Save as RDS
# write_rds(nn_home_act, "Output/nn_home_act.rds")

##### AK #####
# Load data
home_act_initial_ak <- read_xlsx("Input/AK/ak_home_activity_20200113.xlsx")

@@

# Save as RDS
# write_rds(ak_home_act, "Output/ak_home_act.rds")
```

# Join home activity data from different areas
```{r}
wmt_home_act <- read_rds("Output/wmt_home_act.rds")
nn_home_act <- read_rds("Output/nn_home_act.rds")
ak_home_act <- read_rds("Output/ak_home_act.rds")

@@

# write_rds(kids_home_act, "Output/kids_home_act.rds")
```

# Clean up all home activity data
```{r}
kids_home_act <- read_rds("Output/kids_home_act.rds")
kids_linked_ids <- read_rds("Output/kids_linked_ids.rds")

@@

# write_rds(home_act_clean, "Output/home_act_clean.rds")
```

>>>>

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
```

@@

# Load, format, and save home activity data
```{r}
##### WMT #####
# Load data
home_act_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/wmt_home_activity_20200113.csv"))

@@

# Save as RDS
# write_rds(wmt_home_act, paste0(file_path, "Output/wmt_home_act.rds"))

##### NN #####
# Load data
home_act_initial_nn <- read_xlsx(paste0(file_path, "Input/NN/nn_home_activity_20200113.xlsx"))

@@

# Save as RDS
# write_rds(nn_home_act, paste0(file_path, "Output/nn_home_act.rds"))

##### AK #####
# Load data
home_act_initial_ak <- read_xlsx(paste0(file_path, "Input/AK/ak_home_activity_20200113.xlsx"))

@@

# Save as RDS
# write_rds(ak_home_act, paste0(file_path, "Output/ak_home_act.rds"))
```

# Join home activity data from different areas
```{r}
wmt_home_act <- read_rds(paste0(file_path, "Output/wmt_home_act.rds"))
nn_home_act <- read_rds(paste0(file_path, "Output/nn_home_act.rds"))
ak_home_act <- read_rds(paste0(file_path, "Output/ak_home_act.rds"))

@@

# write_rds(kids_home_act, paste0(file_path, "Output/kids_home_act.rds"))
```

# Clean up all home activity data
```{r}
kids_home_act <- read_rds(paste0(file_path, "Output/kids_home_act.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids.rds"))

@@

# write_rds(home_act_clean, paste0(file_path, "Output/home_act_clean.rds"))
```
